A multi-scale story of the diffusion of a new technology: the Web

### 2025-09-28 paper/submission4/tranos2025.qmd
appendix

<!-- main data -->
To capture the diffusion of the commercial Web, a website density metric is
developed for two different geographical scales that are discussed later in 
this section. The main data are counts of commercial websites that are 
calculated using data from the Internet Archive[^2] and, specifically, the 
JISC UK Web Domain Dataset [@ukwebarchive]. The Internet Archive is one of the 
most complete and oldest archive of webpages in the world operating since
1996 [@ainsworth2011much; @holzmann2016dawn]. It is a web crawler, which
discovers webpages by following the hyperlinks of every webpage it
archives. This dataset, which is curated by the British Library,
contains all the archived webpages from the UK country code top level
domain (ccTLD - .uk) from the 1996–2012 period. In essence, this is a
long list of 2.5 billion URLs of archived webpages including also the
archival timestamp.

[^2]: [https://archive.org/](https://archive.org/).

@@

<!-- Multi-scale -->
To empirically test the two hypotheses stated in @sec-litreview, the website 
density metric was calculated for two different scales the Local Authority
Districts (LAD) and the Output Areas (OA). The former is the lowest
administrative unit in the UK and there are 374 LAD. The latter is the smallest
census-based geographical unit and there are c. 230,000 of them in the UK. To 
create the website density metric, the yearly website counts at the LAD level 
are standardised by the number of firms in LAD to avoid biases associated with 
LAD hosting a large number of firms. Given that there is no such readily 
available statistic for the OA, the actual OA level counts are used. However, 
because of the the consistent spatial definition of OA (they host 40-250
households),[^3] website counts in OA are interpreted as a density metric too.

[^3]: <https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies>

@@

<!-- detailed features -->
To capture the hierarchical mechanism the model includes as predictors a
one year lag of website density in London, the largest city in the UK, and a
one year lag of the website density in the nearest city. Both variables clearly
illustrate the urban hierarchy in the UK. Especially for the OA models, a 
third predictor is added to capture hierarchy at a much more granular scale:
a one year lag of the website density in the the nearest retail centre [^4]. As
the literature illustrates, retail centres are key components of urban systems
[@dennis2002central; @jones2021reframing]. Due to the small sizes of the retail 
centres, the latter is only relevant for and included in the OA models and not 
in the LAD models. By including this predictor, the model is better tuned 
towards the spatial characteristic of the OA scale. Hence, it increases the 
robustness of the model.

@@

[^4]: The retail centres data have been obtained by the Consumer Data Research Centre, an ESRC data investment, under project ID CDRC 498-01, ES/L011840/1; ES/L011891/1: <https://data.cdrc.ac.uk/dataset/retail-centre-boundaries-and-open-indicators>.

>>>>

<!-- main data -->
To capture the diffusion of the commercial Web, a website density metric is
developed for two different geographical scales that are discussed later in 
this section. The main data are counts of commercial websites that are 
calculated using data from the Internet Archive[^2] and, specifically, the 
JISC UK Web Domain Dataset [@ukwebarchive][^3]. The Internet Archive is one of the 
most complete and oldest archive of webpages in the world operating since
1996 [@ainsworth2011much; @holzmann2016dawn]. It is a web crawler, which
discovers webpages by following the hyperlinks of every webpage it
archives. This dataset, which is curated by the British Library,
contains all the archived webpages from the UK country code top level
domain (ccTLD - .uk) from the 1996–2012 period. In essence, this is a
long list of 2.5 billion URLs of archived webpages including also the
archival timestamp.

[^2]: [https://archive.org/](https://archive.org/).
[^3]: The data used here is directly available from @tranos_2025_16610235.

@@

<!-- Multi-scale -->
To empirically test the two hypotheses stated in @sec-litreview, the website 
density metric was calculated for two different scales the Local Authority
Districts (LAD) and the Output Areas (OA). The former is the lowest
administrative unit in the UK and there are 374 LAD. The latter is the smallest
census-based geographical unit and there are c. 230,000 of them in the UK. To 
create the website density metric, the yearly website counts at the LAD level 
are standardised by the number of firms in LAD to avoid biases associated with 
LAD hosting a large number of firms. Given that there is no such readily 
available statistic for the OA, the actual OA level counts are used. However, 
because of the the consistent spatial definition of OA (they host 40-250
households),[^4] website counts in OA are interpreted as a density metric too.

[^4]: <https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies>

@@

<!-- detailed features -->
To capture the hierarchical mechanism the model includes as predictors a
one year lag of website density in London, the largest city in the UK, and a
one year lag of the website density in the nearest city. Both variables clearly
illustrate the urban hierarchy in the UK. Especially for the OA models, a 
third predictor is added to capture hierarchy at a much more granular scale:
a one year lag of the website density in the the nearest retail centre [^5]. As
the literature illustrates, retail centres are key components of urban systems
[@dennis2002central; @jones2021reframing]. Due to the small sizes of the retail 
centres, the latter is only relevant for and included in the OA models and not 
in the LAD models. By including this predictor, the model is better tuned 
towards the spatial characteristic of the OA scale. Hence, it increases the 
robustness of the model.

@@

[^5]: The retail centres data have been obtained by the Consumer Data Research Centre, an ESRC data investment, under project ID CDRC 498-01, ES/L011840/1; ES/L011891/1: <https://data.cdrc.ac.uk/dataset/retail-centre-boundaries-and-open-indicators>.

### 2025-09-28 paper/submission4/tranos2025.tex
appendix

To capture the diffusion of the commercial Web, a website density metric
is developed for two different geographical scales that are discussed
later in this section. The main data are counts of commercial websites
that are calculated using data from the Internet Archive\footnote{\url{https://archive.org/}.}
and, specifically, the JISC UK Web Domain Dataset \citep{ukwebarchive}.
The Internet Archive is one of the most complete and oldest archive of
webpages in the world operating since 1996
\citep{ainsworth2011much, holzmann2016dawn}. It is a web crawler, which
discovers webpages by following the hyperlinks of every webpage it
archives. This dataset, which is curated by the British Library,
contains all the archived webpages from the UK country code top level
domain (ccTLD - .uk) from the 1996--2012 period. In essence, this is a
long list of 2.5 billion URLs of archived webpages including also the
archival timestamp.

>>>>

To capture the diffusion of the commercial Web, a website density metric
is developed for two different geographical scales that are discussed
later in this section. The main data are counts of commercial websites
that are calculated using data from the Internet Archive\footnote{\url{https://archive.org/}.}
and, specifically, the JISC UK Web Domain Dataset
\citep{ukwebarchive}\footnote{The data used here is directly available
  from \citet{tranos_2025_16610235}.}. The Internet Archive is one of
the most complete and oldest archive of webpages in the world operating
since 1996 \citep{ainsworth2011much, holzmann2016dawn}. It is a web
crawler, which discovers webpages by following the hyperlinks of every
webpage it archives. This dataset, which is curated by the British
Library, contains all the archived webpages from the UK country code top
level domain (ccTLD - .uk) from the 1996--2012 period. In essence, this
is a long list of 2.5 billion URLs of archived webpages including also
the archival timestamp.
